Supporting Information : Impact of asymmetric tail-dependence in extreme climatic events on extinction risk in population ecology

### 2019-04-06 MainText.Rmd
started writing results

To test our hypothesis from (Q1), we repeated the above analysis 
considering the vital rates coming from a 
an asymmetric copula with specific 
left (Clayton, Survival Gumbel)
or right (Survival Clayton, Gumbel) tail dependence
or from an extreme tail-depentdent one, 
when vital rates were considered to be affected
by extreme climatic events. A $N \times 2$ matrix, M, was generated
using `BiCopSim` function
from `VineCopula` package 
for each of the left 
(Clayton, Survival Gumbel) and right 
(Survival Clayton, Gumbel) tail-dependent bivariate copulas, 
where $N = numsteps(=25) \times numsims(=10000)$.
We kept the same Kendall's correlation ($\tau = 0.6$) for 
each bivariate copula from one parameter family which used its parameter accordingly
(see `iTau` function from `copula` package). Entries within each clumn of
matrix $M$, were thus 
independent of each others but 
the two columns of $M$ had a specified Kendall's 
correlation ($\tau = 0.6$) and underlying tail-dependence. 
Similarly, for extreme-tail dependent bivariant copula
(either left or right tail) a matrix M was generated from standard iid as described in 
Appendix \ref{SM-extr_tail_dep}.

@@

<!--Q1 : age structured results : linear and nonlinear model with asymmetric copula : Fig. S1-->

<!--Q1 : age structured results : linear and nonlinear model with extreme tail dep. copula : Fig. 1 -->

@@

 <!--FIG : age structure model results : linear, nonlinear-->
\begin{figure*}[!h]
\textbf{ \hspace{5.4 cm} (A) \hspace{4 cm} (B)} \\
\vspace{-0.9 cm}
\begin{center}
\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_extremetail_linear.pdf}
\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_extremetail_nonlinear_omg_100.pdf}\\
\caption{Extinction risk from single patch age structured $(A)$ linear ($\Omega = \infty$) and $(B)$ nonlinear ($\Omega = 100$) models, fecundity of adults and survival of eggs have either extreme left or extreme right tail dependence, numsims=10000. \label{fig_extrisk_agestr}}
\end{center}
\end{figure*}

>>>>

To test our hypothesis from (Q1), we repeated the above analysis 
considering the vital rates coming from an asymmetric copula with specific 
left (Clayton, Survival Gumbel)
or right (Survival Clayton, Gumbel) tail dependence
or from an extreme tail-depentdent one, 
when vital rates were considered to be affected
by extreme climatic events. A $N \times 2$ matrix, M, was generated
using `BiCopSim` function
from `VineCopula` package 
for each of the left 
(Clayton, Survival Gumbel) and right 
(Survival Clayton, Gumbel) tail-dependent bivariate copulas, 
where $N = numsteps(=25) \times numsims(=10000)$.
We kept the same Kendall's correlation ($\tau = 0.6$) for 
each bivariate copula from one parameter family which used its parameter accordingly
(see `iTau` function from `copula` package). Entries within each clumn of
matrix $M$, were thus 
independent of each others but 
the two columns of $M$ had a specified Kendall's 
correlation ($\tau = 0.6$) and underlying tail-dependence. 
Similarly, for extreme-tail dependent bivariant copula
(either left or right tail) a matrix M was generated from standard iid as described in 
Appendix \ref{SM-extr_tail_dep}.

@@

<!--Q1 : age structured results : linear and nonlinear model with asymmetric copula : Fig. S1-->
<!--Q1 : age structured results : linear and nonlinear model with extreme tail dep. copula : Fig. 1 -->

For single patch linear age-structured models, we found left-tail 
dependence (Clayton, Suvival Gumbel copulas) in noisy vital 
rates enhanced the extinction risk compared to the case when
vital rates were right tail dependent (Survival clayton, Gumbel copulas).
It was expected because a sudden "bad" year (e.g. with severe cold) can suppress 
the fecundity of adult and survival/ maturation of eggs simultaneously 
and thereby increasing the extinction risk of the population. For non-linear age-structured 
models, we got the similar trend for extinction risks with different tail-dependence (left and right).
With non-linear model the extinction risks appeared as higher than their corresponding 
linear analog as it was evident from Eq. \ref{eq_vitalrates} with finite $\Omega = 100$. 
Fig. \ref{SM-fig_extrisk_agestr_C_SC_G_SG_linear} represents this scenario with 4 different 
asymmetric copulas but of same Kendall's correlation ($\tau = 0.6$). There was very little
difference between extinction risks within a pair of different tail-dependent copulas
(e.g. C-SC in Fig. \ref{SM-fig_extrisk_agestr_C_SC_G_SG_linear}(A,C) and
G-SG in Fig. \ref{SM-fig_extrisk_agestr_C_SC_G_SG_linear}(B,D)) and that difference was partly 
dependent on the strength of tail dependence. Therefore we repeated the same analysis with 
vital rates coming from extreme-tail dependent (left and right in two different simulations)
copulas and got the similar results but with much prominent difference between their extinction risks
(Fig. \ref{fig_extrisk_agestr}). This result answered (Q1) and showed extinction risk of 
a population could be considered as
direct manifestation of tail-dependence measure between vital rates of different age-class
within that population. <!--discussion later : add Tuljapurkar's approximation-->

@@

 <!--FIG : age structure model results : linear, nonlinear-->
\begin{figure*}[!h]
\textbf{ \hspace{5.4 cm} (A) \hspace{4 cm} (B)} \\
\vspace{-0.9 cm}
\begin{center}
\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_extremetail_linear.pdf}
\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_extremetail_nonlinear_omg_100.pdf}\\
\caption{Extinction risk from single patch age structured $(A)$ linear ($\Omega = \infty$) and $(B)$ nonlinear ($\Omega = 100$) models
against time (shading represents 95\% confidence interval). Fecundity of adults and survival of eggs have either extreme left or extreme right tail dependence, numsims=10000. \label{fig_extrisk_agestr}}
\end{center}
\end{figure*}

### 2019-04-06 SuppMat.Rmd
started writing results

<!--Supp. FIG :  age structure model results for C, SC, G, SG with linear model (\Omega = infinity)-->
\begin{figure}[!h]
\begin{center}
\textbf{ \hspace{0.5 cm} (A) \hspace{4.4 cm} (B)} \\
\vspace{-0.2 cm}
\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_C_SC_linear.pdf}
\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_SG_G_linear.pdf}\\
\textbf{ \hspace{0.5 cm} (C) \hspace{4.4 cm} (D)} \\
\vspace{-0.2 cm}
\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_C_SC_nonlinear_omg_100.pdf}
\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_SG_G_nonlinear_omg_100.pdf}\\
\caption{Extinction risk from single patch age structured models, fecundity of adults and survival of eggs have either left or right tail dependence in their corresponding copula structure $(A)$ Clayton and Survival Clayton (parameter = 3), $(B)$ Survival Gumbel and Gumbel (parameter = 2.5) for linear model ($\Omega = \infty$), numsims=10000. $(C - D)$ represents the analogous figures as of $(A - B)$ but for non-linear model ($\Omega = 100$). All copula has same Kendall correlation $\tau = 0.6$. \label{fig_extrisk_agestr_C_SC_G_SG_linear}}
\end{center}
\end{figure}

>>>>

<!--Supp. FIG :  age structure model results for C, SC, G, SG with linear model (\Omega = infinity)-->
\begin{figure}[!h]
\begin{center}
\textbf{ \hspace{0.5 cm} (A) \hspace{4.4 cm} (B)} \\
\vspace{-0.2 cm}
\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_C_SC_linear.pdf}
\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_SG_G_linear.pdf}\\
\textbf{ \hspace{0.5 cm} (C) \hspace{4.4 cm} (D)} \\
\vspace{-0.2 cm}
\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_C_SC_nonlinear_omg_100.pdf}
\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_SG_G_nonlinear_omg_100.pdf}\\
\caption{Extinction risk (shading represents 95\% confidence interval) from single patch age structured models with time; fecundity of adults and survival of eggs have either left or right tail dependence in their corresponding copula structure. $(A)$ Clayton and Survival Clayton (parameter = 3), $(B)$ Survival Gumbel and Gumbel (parameter = 2.5) for linear model ($\Omega = \infty$), numsims=10000. $(C - D)$ represents the analogous figures as of $(A - B)$ but for non-linear model ($\Omega = 100$). All copula has same Kendall correlation $\tau = 0.6$. \label{fig_extrisk_agestr_C_SC_G_SG_linear}}
\end{center}
\end{figure}
